---
title: 'Birth Prevalence of Neural Tube Defects in East Africa: A Systematic Review and Meta-Analysis'
author: 'Paddy Ssentongo MD, PhD'
date: "02/22/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


# LOAD REQUIRED PACKAGES AND FUNCTIONS -----------------------------------------
```{r, message=FALSE, echo=FALSE,warning=FALSE}

library(ggplot2) 
library(metafor) 
library(meta) 
library(tidyr)
library(dplyr)
library(ggrepel)
library(ggpmisc)

```


```{r setup, include=FALSE,warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning =F)
```



# read data -----------------------------------------
```{r,echo=FALSE,warning=FALSE}
rm(list=ls())
#data=read.csv("Subdural_Lit_Rev.csv")
dat=read.csv("data/NTD_Meta77.csv")
```


# Meta-analysis using DL and logit transformation-----------------------------------------
```{r,echo=FALSE,warning=FALSE}
ntd <- metaprop(NTD, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP", hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")

sb <- metaprop(spina.bifida, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP",  hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")

enc<- metaprop(Encephalocele, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP",  hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")

ane<- metaprop(anencephaly, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP",  hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")

```


#Influence analysis  using leave-one-out -----------------------------------------
```{r, message=F,echo=FALSE,fig.height=5, fig.width=6, echo=FALSE,warning=FALSE}
pdf("figs/SenstivityPlots.pdf", width = 10, height = 8)
forest(metainf(ntd, pooled = "random"))
dev.off()

inf1=metainf(ntd, pooled = "random")# ntd
inf2=metainf(sb, pooled = "random")# sb
inf3=metainf(enc, pooled = "random") # enc
inf4=metainf(ane, pooled = "random") # ane


```
#Influence analysis  using leave-one-out NTD-----------------------------------------
```{r, message=F,echo=FALSE,fig.height=7, fig.width=10, echo=FALSE,warning=FALSE}
forest(inf1,
       leftlabs = c("Author"),
       digits = 2,
       col.by="black",
       #sortvar = "TE",
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Birth Prevalence of Neural Tube Defects in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(20, 60), at = c(20,30,40,50,60),pscale = 10000,
       backtransf = T)

```


#Influence analysis  using leave-one-out SB-----------------------------------------

```{r, message=F,echo=FALSE,fig.height=7, fig.width=10, echo=FALSE,warning=FALSE}
forest(inf2,
       leftlabs = c("Author"),
       digits = 2,
       col.by="black",
       #sortvar = "TE",
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Birth Prevalence of Spina Bifida in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(10, 40), at = c(10,20,30,40),pscale = 10000,
       backtransf = T)

```


#Influence analysis  using leave-one-out EC-----------------------------------------


```{r, message=F,echo=FALSE,fig.height=8, fig.width=10, echo=FALSE,warning=FALSE}
forest(inf3,
       leftlabs = c("Author"),
       digits = 2,
       col.by="black",
       #sortvar = "TE",
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Birth Prevalence of Encephalocele in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(1, 5), at = c(1,2,3,4),pscale = 10000,
       backtransf = T)

```
#Influence analysis  using leave-one-out AN-----------------------------------------


```{r, message=F,echo=FALSE,fig.height=7, fig.width=10, echo=FALSE,warning=FALSE}
forest(inf4,
       leftlabs = c("Author"),
       digits = 2,
       col.by="black",
       #sortvar = "TE",
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
       xlab="Birth Prevalence of Anacephaly in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(3, 14), at = c(3,5,10,14),pscale = 10000,
       backtransf = T)

```


# Funnel plots : Prevalence -----------------------------------------
```{r}

pdf("figs/FunnelPlotFinalNTD.pdf", width = 8, height = 7)
funnel(ntd,xlab = "log transformed proportions",studlab = F)
dev.off()

metabias(ntd, method="linreg")

```

# trim and fill : Incidence -----------------------------------------
```{r,fig.height=5, fig.width=4, echo=FALSE,warning=FALSE}
tf2 <- trimfill(ntd)
summary(tf2)
pdf("figs/Trimmed.pdf", width = 7, height = 7)
funnel(tf2)
dev.off()
#funnel(tf2, pch = ifelse(tf2$trimfill, 1, 16),
      # level = 0.90, comb.random = T,backtransf = T)
#
```

```{r}
tf3 <- trimfill(ntd)
summary(ntd)

```



# Funnel plots : Incidence -----------------------------------------
```{r,echo=FALSE,warning=FALSE}

pdf("figs/FunnelPlot.pdf", width = 10, height = 10)
funnel(ntd,xlab = "log transformed proportions",studlab = TRUE)
dev.off()

metabias(ntd, method="linreg")# Egger's 
metabias(ntd, method="rank")# Begg's
```

# Descriptive trend
```{r}
dat=read.csv("data/NTD_Meta77.csv")
formula <- y ~ x

f <- dat %>%
  arrange(desc(N)) %>%
  ggplot(aes(x = Cohort_Year, y = rate_10000)) +
  #ggplot(aes(x = Period, y = Rates, fill=Period))+  
  #geom_smooth(se = F, method = "lm", size = 1.5) +
  #geom_boxplot(show.legend = F)+
  geom_point(aes(fill = Country,  size = N), shape = 21, alpha = .7) +
    geom_smooth(method=lm , color="red", se=F, size=1.5, lty=1) +
  colorblindr::scale_fill_OkabeIto(name = "Eastern Africa \n Country (UN)") +
  scale_size(name = "Sample Size",
             range = c(2, 10), breaks = c(10, 50)) +
  annotate("text", label = "R^{2}==0.11~~p==0.08", x = 1989, y = 130, parse = T) +
  geom_text_repel(aes(label = Country), size = 2.5) +
  #expand_limits(x=c(1983,2018))+
  scale_x_continuous(breaks=seq(1983,2018,5))+
  scale_y_continuous(breaks=seq(0,120,10))+
  facet_wrap(~ "Year of study") +
  labs(x = "Year of study", y = "NTD Rates (per 10,000)")  +
  guides(fill = guide_legend(override.aes = list(size = 3.5))) +
  theme_bw(base_size = 14) +
  theme(strip.text = element_text(face = "bold", size = 13))



ggpubr::ggarrange(f, ncol = 1, common.legend = T, legend = "top")
ggsave("figs/Fig5.pdf", width = 10, height = 8)

dat$YEAR=dat$Cohort_Year
model_A <- rma.glmm( xi=NTD, ni=N, measure="PLO",  mods = ~ YEAR, dat=dat)
print(model_A,digits=3)
model_A 
round(exp(coef(summary(model_A))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)



```


```{r}
dat=read.csv("data/NTD_Meta77.csv")
dat3 <- escalc(measure="PLO", xi=NTD, ni=N,data=dat)
### mixed-effects meta-regression model 
res <- rma(yi, vi, mods = ~ Cohort_Year, data=dat3)
print(res, digits=3)
round(exp(coef(summary(res))[-1,c("estimate", "ci.lb", "ci.ub")]), 2)


```



# Plot figure 2: Overall
```{r,fig.height=9, fig.width=10, echo=FALSE,warning=FALSE}
pdf("figs/Fig2.pdf", width = 9, height = 8)
forest(ntd,
       leftlabs = c("Author, Year", "Cases", "Sample size"),
       digits = 2,
       col.by="black",
       #sortvar = Year_Publication,
       sortvar = TE,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
        xlab="Birth Prevalence of Neural Tube Defects in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0,70), at = c(0,10,20,30,40,50,60,70),pscale = 10000,
       backtransf = T)
dev.off()

```


# Plot figure 4: Country
```{r,fig.height=15, fig.width=10, echo=FALSE,warning=FALSE}
mw<- update(ntd, byvar=dat$Country,print.byvar=F)
pdf("figs/Fig3.pdf", width = 10, height = 12)
forest(mw,
       leftlabs = c("Author, Year", "Cases", "Sample size"),
       digits = 2,
       col.by="black",
       sortvar = Year_Publication,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
         xlab="Birth Prevalence of Neural Tube Defects in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0,70), at = c(0,10,20,30,40,50,60,70),pscale = 10000,
       backtransf = T)
dev.off()

```







# Plot figure 3A: SB
```{r,fig.height=9, fig.width=10, echo=FALSE,warning=FALSE}
pdf("figs/fig4A.pdf", width = 9, height = 8)
forest(sb,
       leftlabs = c("Author, Year", "Cases", "Sample size"),
       digits = 2,
       col.by="black",
       #sortvar = Year_Publication,
       sortvar = TE,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
        xlab="Birth Prevalence of Spina Bifida in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0,70), at = c(0,10,20,30,40,50,60,70),pscale = 10000,
       backtransf = T)
dev.off()

```



# Plot figure 3B: ANE
```{r,fig.height=9, fig.width=10, echo=FALSE,warning=FALSE}
dat=read.csv("data/NTD_Meta77.csv")

l1=is.na(dat$anencephaly)
dat=dat[!l1,]

ane<- metaprop(anencephaly, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP",  hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")


pdf("figs/Fig4B.pdf", width = 9, height = 7)
forest(ane,
       leftlabs = c("Author, Year", "Cases", "Sample size"),
       digits = 2,
       col.by="black",
       #sortvar = Year_Publication,
       sortvar = TE,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
        xlab="Birth Prevalence of Anencephaly in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0,70), at = c(0,10,20,30,40,50,60,70),pscale = 10000,
       backtransf = T)
dev.off()

```


# Plot figure 3C: ENCEPHALOCELE
```{r,fig.height=9, fig.width=10, echo=FALSE,warning=FALSE}
dat=read.csv("data/NTD_Meta77.csv")

l1=is.na(dat$Encephalocele)
dat=dat[!l1,]

enc<- metaprop(Encephalocele, N, data = dat, studlab = paste(Author, Year_Publication,  sep = ", "), pscale = 10000, 
               comb.fixed = F, method.ci = "CP",  hakn=TRUE,
               method.tau = "DL", method = "Inverse",sm = "PLOGIT")
               #method.tau = "ML", method = "GLMM",sm = "PLOGIT")


pdf("figs/Fig4C.pdf", width = 9, height = 5)
forest(enc,
       leftlabs = c("Author, Year", "Cases", "Sample size"),
       digits = 2,
       col.by="black",
       #sortvar = Year_Publication,
       sortvar = TE,
       squaresize=0.5,
       col.square="blue",
       col.diamond="maroon",
       col.diamond.lines="maroon",
        xlab="Birth Prevalence of Encephalocele in East Africa (per 10,000 population)",
       print.pval.Q = T,
       digits.pval.Q=4,
       print.tau2=F,
       xlim = c(0,70), at = c(0,10,20,30,40,50,60,70),pscale = 10000,
       backtransf = T)
dev.off()

```






